Simulations of Ballot Polling Risk-Limiting Audits

### 2022-10-11 providence/sections/prov.tex

\begin{definition}[$(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence]
    \label{def:minervatwo}
    For cumulative round size $n_i$ for round $i$ and a cumulative $k_i$ ballots for the reported winner found in round $i$, the \R \Providence stopping rule for the $j^{th}$ round is:
$$
\mathcal{A}(X_{j})=  \left\{ \begin{array}{ll} \text{Correct} ~~~~ \omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_j, n_{j-1}) \geq \frac{1}{\alpha}\\
        % & \\
        % incorrect& ~~~ \sigma_n < \frac{\beta}{1-\alpha} \\
        % & \\
        Undetermined ~~else \\
    \end{array}
    \right .
$$
where $\omega _{1}\triangleq \tau_{1}$ and for $j\ge 2$, we define $\omega _{j}$ as follows:
\begin{equation}
    \begin{aligned}
    \omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_{j}, n_{j-1})
    \triangleq\\
    \sigma(k_{j-1},p_a,p_0,n_{j-1})\cdot \tau_1(k_{j}-k_{j-1},p_a,p_0,n_j-n_{j-1})
    \end{aligned}
\end{equation}
\end{definition}

Notice that for $j\ge 2$, unlike $\tau_j$, computing $\omega_j$ requires no convolution. We now we prove that \Providence is risk-limiting using lemmas from basic algebra in Appendix~\ref{sec:proofs}.

@@

For any round $j\ge 2$, by Definition \ref{def:minervatwo}
and Lemma \ref{lemma:minerva2_kmin_exists},
$\mathcal{A}=\text{Correct}$ (the audit stops) only when
$$
\omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_{j}, n_{j-1}, \alpha )\triangleq
\sigma(k_{j-1},p_a,p_0,n_{j-1})\cdot \tau_1(k_{j}-k_{j-1},p_a,p_0,n_j-n_{j-1})
\ge \frac{1}{\alpha}.
$$
%Let $d_j=k_j-k_{j-1}$ be shorthand for the new draw.
%could do this for simpler notation^
By Lemma \ref{lemma:any_ratio_is_sigma_simple}
and Definition \ref{def:minerva_ratio}, this is equivalent to
$$
\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge \frac{1}{\alpha}.
$$
By Lemma~\ref{lemma:minerva2_kmin_exists} and Definition~\ref{def:minervatwo},
we see that there exists a $k_{min, j}\le k_j$ for which
$$
\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge
$$
$$
\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_0, n{j}]} \ge 
$$
$$
\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge \frac{1}{\alpha}.
$$

@@

\subsection{Resistance against an adversary choosing round sizes}
I don't feel very comfortable with how to word/present this argument... I have not yet made an attempt at doing so.

>>>>

In this section we motivate and introduce the stopping condition of \Providence.

The \Minerva stopping condition tests the tail ratios of the probability distributions (given the null and alternate hypotheses) of winner votes $K_j$ in round $j$. Note that these distributions are not conditioned on $k_{j-1}$. Thus they may be viewed as a weighted average of various distributions, each conditioned on a possible value of $K_{j-1}$. The proof of \Minerva's risk-limiting property relies on the fact that, at every stopping point, the tail of the probability distribution function given $H_0$ is at most $\alpha$ times that given $H_a$. However, because these distributions are averaged over the various values of $K_{j-1}$, this approach to the proof works only if the next round size is the same for all values of $K_{j-1}$. If not, this average pdf will not correctly represent the pdf of $K_j$. For example, if some values of $K_{j-1}$ would result in larger next round sizes, the pdf for these values of $K_{j-1}$ should not contribute to the average distribution function of $K_j$ for smaller round sizes.

For the \Providence audit we test the tail ratios of the distributions conditioned on $k_{j-1}$ and multiplied by the weighting factor. That is, we test this tail ratio separately for each weighted distribution corresponding to each possible value of $K_{j-1}$. Because of this, we no longer require that all values of $k_{j-1}$ should result in the same next round size. This results in a great deal of flexibility in multiple round audits.

\subsection{Definition}
\label{sec:prov_def}
\begin{definition}[$(\alpha,p_a, p_0,k_{j-1},n_{j-1},n_j)$-\Providence]
    \label{def:minervatwo}
    For cumulative round size $n_i$ for round $i$ and a cumulative $k_i$ ballots for the reported winner found in round $i$, the \R \Providence stopping rule for the $j^{th}$ round is:
$$
\mathcal{A}(X_{j})=  \left\{ \begin{array}{ll} \text{Correct} ~~~~ \omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_j, n_{j-1}) \geq \frac{1}{\alpha}\\
        % & \\
        % incorrect& ~~~ \sigma_n < \frac{\beta}{1-\alpha} \\
        % & \\
        Undetermined ~~else \\
    \end{array}
    \right .
$$
where $\omega _{1}\triangleq \tau_{1}$ and for $j\ge 2$, we define $\omega _{j}$ as follows:
\begin{equation}
    \begin{aligned}
    \omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_{j}, n_{j-1})
    \triangleq\\
    \sigma(k_{j-1},p_a,p_0,n_{j-1})\cdot \tau_1(k_{j}-k_{j-1},p_a,p_0,n_j-n_{j-1})
    \end{aligned}
\end{equation}
\end{definition}

Notice that for $j\ge 2$, unlike $\tau_j$, computing $\omega_j$ requires no convolution and is hence considerably more computationally efficient.

\subsection{Proof of Risk-Limiting Property}
\label{sec:proof}
We now prove that \Providence is risk-limiting using lemmas from basic algebra in Appendix~\ref{sec:proofs}.

@@

For any round $j\ge 2$, by Definition \ref{def:minervatwo}
and Lemma \ref{lemma:minerva2_kmin_exists},
$\mathcal{A}=\text{Correct}$ (the audit stops) only when
\begin{equation*}
\begin{aligned}
\omega_{j}(k_{j}, k_{j-1}, p_a, p_0, n_{j}, n_{j-1}, \alpha )\triangleq\\
\sigma(k_{j-1},p_a,p_0,n_{j-1})\cdot \tau_1(k_{j}-k_{j-1},p_a,p_0,n_j-n_{j-1})
\ge \frac{1}{\alpha}.
\end{aligned}
\end{equation*}
%Let $d_j=k_j-k_{j-1}$ be shorthand for the new draw.
%could do this for simpler notation^
By Lemma \ref{lemma:any_ratio_is_sigma_simple}
and Definition \ref{def:minerva_ratio}, this is equivalent to
$$
\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge \frac{1}{\alpha}.
$$
By Lemma~\ref{lemma:minerva2_kmin_exists} and Definition~\ref{def:minervatwo},
we see that there exists a $k_{min, j}\le k_j$ for which
$$
\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge
$$
$$
\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{min, j} \mid \bm{k_{j-1}}, H_0, n{j}]} \ge 
$$
$$
\frac{\Pr[\bm{k_{j-1}} \mid H_a]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_a, n{j}]}{\Pr[\bm{k_{j-1}} \mid H_0]\cdot Pr[K_{j} \ge k_{j} \mid \bm{k_{j-1}}, H_0, n{j}]}\ge \frac{1}{\alpha}.
$$

@@

\subsection{Resistance to an adversary choosing round sizes}
\label{sec:adversary}
Filip TBD
